---
title: "nonmem2mrgsolve"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, eval=FALSE) # echo = FALSE
```

## Purpose

Provide an R package that automates NONMEM to mrgsolve translation to reduce human errors and improve efficiency.

## Introduction

Pharmacometricians often utilize multiple software, with NONMEM (ICON plc, Dublin, Ireland) widely regarded as the current gold standard for population pharmacokinetic (PK) and pharmacokinetic-pharmacodynamic (PK-PD) modeling. While NONMEM has the ability to simulate, it is often more convenient and practical to perform simulations using free open-source software such as R which can also perform post-processing of the output.A popular pharmacometrics-orientated R package for solving ODE systems is mrgsolve. Although mrgsolve has built-in functionality to streamline translation, to the best of our knowledge a previously-developed and publicly-available R package for fully automated NONMEM to mrgsolve translation does not exist.

## Installation

If not already installed, install R. Then within R run the following command:
```{r}
devtools::install_github("Andy00000000000/nonmem2mrgsolve")
```

## Getting Started

Full work-through examples will be created shortly, but in the meantime assume that a NONMEM run has finished estimating and has explicitly-defined ODEs. The ctl and ext files are both named pbpk-101 and are located at C:/Documents/NONMEM/Project.

Load R and run:
```{r}
setwd("C:/Documents/NONMEM/Project") # set the working directory to where the NONMEM files are located
library(nonmem2mrgsolve)             # load the nonmem2mrgsolve package
nonmem2mrgsolve("pbpk-101")          # run the nonmem2mrgsolve function with default settings
```

The resulting mrgsolve code will be wrote to mrgsolve-code-V0_pbpk-101.R within the C:/Documents/NONMEM/Project folder. The user should then validate the translation, for which there is an intuitive and established framework [(Here)](https://mrgsolve.org/blog/posts/2023-update-validation.html).

## Usage Statement

```{r}
nonmem2mrgsolve::nonmem2mrgsolve(
  filename = "nonmem-model.ctl", # String of the NONMEM model name with or without the .ctl extension
  dir = "./folder",              # String of the directory path to the NONMEM files (if not already given in the file name input or the working directory already set)
  sigdig = NULL,                 # Numeric of the number of significant digits to round non-fixed thetas and etas to; default NULL for no rounding
  write = T,                     # [T or F] for whether to write the resulting mrgsolve code to an R file
  return.orig = F,               # [T or F] for whether to return the original NONMEM ctl and ext (or cnv) files to the R environment
  out.filename = NULL,           # String of the name for the mrgsolve output file without the .R extension; NULL for default naming
  use.cnv = F                    # [T or F] for whether to use the NONMEM cnv file for parameter estimates instead of the ext file
)
```

## Getting Help

Please feel free to report problems using the [Issue Tracker](https://github.com/Andy00000000000/nonmem2mrgsolve/issues) or to reach out for help on the [Discussion Board](https://github.com/Andy00000000000/nonmem2mrgsolve/discussions).

## Our Team

[Enhanced Pharmacodynamics, LLC (ePD)](https://www.epd-llc.com/) is a contract research organization that assists clients with the design and implementation of model-informed drug development strategies in a broad range of therapeutic areas. The executive management team is led by [Dr. Donald E. Mager](https://www.linkedin.com/in/don-mager-2652615/) and [Dr. Scott A. Van Wart](https://www.linkedin.com/in/scott-van-wart-2b6002a/).

[Finch Studio](https://finchstudio.io/) is an integrated modeling environment for pharmacometricians, clinical pharmacologists, and other team members to visualize, develop, and organize PK/PD models and data. It includes a modern NONMEM code editor and workbench, interactive data tables and visualizations, and a PK/PD model library. The development team is led by [Dr. Mohamed Ismail](https://www.linkedin.com/in/mohamed-ismail-bb54a367/).

nonmem2mrgsolve was developed at ePD by [Andrew Santulli](https://www.linkedin.com/in/andrew-santulli-219034156/).
